---
title: "ElkhornSloughDO"
author: "Juliana Cornett"
date: "1/10/2022"
output: html_document
---

#Plot of DO in Elkhorn Slough

#install packages

```{r}
#install.packages("tidyr")
#install.packages("ggplot2")
#install.packages("tidyverse")
#install.packages("ggpubr")
#install.packages("lubridate")
#install.packages("scales")
#install.packages("viridis")
#install.packages("RColorBrewer")
#install.packages("patchwork")
```

#load libraries

```{r}
library(tidyr)
library(ggplot2)
library(tidyverse)
library(ggpubr)
library(lubridate)
library(scales)
library(viridis)
library(RColorBrewer)
library(patchwork)
```

#read in csv files containing DO measurements from 5/1/31-8/31/21 at the 4 water quality sites in Elkhorn Slough (Vierra Mouth, South Marsh, North Marsh, and Azevedo Pond)

```{r}
Vierra <- read.csv("VierraMouth_DO_5.1.21_8.31.21.csv")

South <- read.csv("SouthMarsh_DO_5.1.21_8.31.21.csv")

North <- read.csv("NorthMarsh_DO_5.1.21_8.31.21.csv")

Azevedo <- read.csv("AzevedoPond_DO_5.1.21_8.31.21.csv")

vm_temp <- read.csv("ELKVMWQ.csv") %>% dplyr::select(Station_Code, DateTimeStamp, Temp)

sm_temp <- read.csv("ELKSMWQ.csv") %>% dplyr::select(Station_Code, DateTimeStamp, Temp)

nm_temp <- read.csv("ELKNMWQ.csv") %>% dplyr::select(Station_Code, DateTimeStamp, Temp)

ap_temp <- read.csv("ELKAPWQ.csv") %>% dplyr::select(Station_Code, DateTimeStamp, Temp)
```

#combine into single dataframe

```{r}
AllES_DO <- rbind(Vierra, South, North, Azevedo)

All_Temp <- rbind(vm_temp, sm_temp, nm_temp, ap_temp)
```

#remove rows with NA

```{r}
ES_DO <- na.omit(AllES_DO)

All_Temp <- na.omit(All_Temp)
```

#fix timestamp column
#first change slashes in year of TimeStamp column to dashes with gsub

```{r}
ES_DO$TimeStamp <- gsub('/', '-', ES_DO$TimeStamp)
ES_DO

All_Temp$TimeStamp <- gsub('/', '-', All_Temp$DateTimeStamp)
All_Temp
```

#then convert time stamp into POSIXct format, specifying the format that the timestamp is currently in

```{r}
ES_DO$TimeStamp <- as.POSIXct(ES_DO$TimeStamp, format="%m-%d-%y %H:%M")
ES_DO$Date <- date(ES_DO$TimeStamp)
ES_DO

All_Temp$TimeStamp <- as.POSIXct(All_Temp$TimeStamp, format="%m-%d-%Y %H:%M")
All_Temp$Date <- date(All_Temp$TimeStamp)
All_Temp
```

#plot to visualize across entire summer

```{r}
ggplot(ES_DO, aes(x = TimeStamp, y = DO_mgL, color = Site)) + 
  theme_pubr() + 
  geom_line() +
  ylab("Dissolved Oxygen (mg/L)") + 
  xlab("Date") + 
  theme(axis.text.x = element_text(angle = 90)) + 
  theme(axis.title=element_text(size=18,face="bold")) +
  theme(axis.text=element_text(size=10)) + 
  theme(legend.text=element_text(size=12,face="bold")) +
  theme(legend.title = element_blank()) 
```

#add month column

```{r}
ES_DO$Month <- as.numeric(format(ES_DO$TimeStamp, "%m"))
ES_DO
```

#subset May data

```{r}
May_DO <- subset(ES_DO, Month==5)
May_DO
```

#set my own color palette based on colors chosen from viridis palette

```{r}
palette_viridis <- c("#482677FF", "#2D7D8EFF", "#55C667FF","#FDE725FF")
```

#plot May only (Figure 1 in manuscript)

```{r}
es_do_may <- ggplot(May_DO, aes(x = TimeStamp, y = DO_mgL, color = Site)) + 
  scale_color_manual(values = palette_viridis) +
  theme_pubr() + 
  geom_line() +
  guides(colour = guide_legend(override.aes = list(size=3))) +
  ylab("Dissolved Oxygen (mg/L)") + 
  xlab("") + 
  theme(axis.text.x = element_text(angle = 90)) + 
  theme(axis.title=element_text(size=14,face="bold")) +
  theme(axis.text=element_text(size=10)) + 
  theme(legend.text=element_text(size=10,face="bold")) +
  theme(legend.title = element_blank()) +
  scale_x_datetime(date_breaks = "1 day", date_labels = "%b %d") + 
  geom_hline(yintercept=8, linetype='dotted') +
  geom_text(data=May_DO[1, ], label="A", vjust=-12, color = "black", size = 5, check_overlap = TRUE) +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
es_do_may

es_temp_may <- ggplot(All_Temp, aes(x = TimeStamp, y = Temp, color = Station_Code)) + 
  scale_color_manual(values = palette_viridis) +
  theme_pubr() + 
  geom_line() +
  guides(colour = guide_legend(override.aes = list(size=3))) +
  ylab("Temperature (\u00B0C)") + 
  xlab("Date") + 
  theme(axis.text.x = element_text(angle = 90)) + 
  theme(axis.title=element_text(size=14,face="bold")) +
  theme(axis.text=element_text(size=10)) + 
  theme(legend.text=element_text(size=10,face="bold")) +
  theme(legend.title = element_blank()) +
  scale_x_datetime(date_breaks = "1 day", date_labels = "%b %d") + 
  geom_hline(yintercept=12, linetype='dotted') +
  geom_text(data=All_Temp[1, ], label="B", vjust=-21, color = "black", size = 5, check_overlap = TRUE)+
  theme(legend.position = "none")
es_temp_may
```

combined jpeg

```{r}
jpeg("ElkhornSlough_TempDO.jpeg", width = 8, height = 10, units = "in", res = 600)
es_do_may /
  es_temp_may
```

